---
title: "11-assignment_Hurst.Rmd"
author: "Jason Hurst"
output: github_document
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Get libraries
```{r}
library(tidyverse)
library(nycflights13) 
# `nycflights13` database, which contains information on every single flight departing New York City in 2013, including airline information, airport information, flight information, plane information and weather information. 
library(RSQLite)
library("viridis")
library("datamodelr")
```

## 1. Using the nyclfights13 library, open all of the data tables and turn them into a database. Make sure to include flights, airlines, airports, weather and planes.

# NYC Flights

Open tables. They're stored in `data()` from the `nycflights` library. 

```{r}
# Data sources from NYC flights
data(flights)
data(airlines)
data(airports)
data(weather)
data(planes)
```

# Connecting to a database

`con` below refers to our connection to a database. It remains open until we close it. It will be based on the properties of the database we're trying to access. 

```{r}
con <- dbConnect(RSQLite::SQLite(), ":memory:")

#Write flights tables to database (you won't usually do this)
dbWriteTable(con,
             "flights", 
             as.data.frame(flights))

dbWriteTable(con,
             "planes", 
             as.data.frame(planes))

dbWriteTable(con,
             "airlines", 
             as.data.frame(airlines))

dbWriteTable(con,
             "weather", 
             as.data.frame(weather))

dbWriteTable(con,
             "airports", 
             as.data.frame(airports))
```


## Tables and Fields
Check access to tables.
```{r}
#List tables in database
dbListTables(con)
```

For some of these tables, what fields are included? 

```{r}
#List fields in various tables
dbListFields(con,"airlines")

dbListFields(con,"flights")

dbListFields(con,"planes")

dbListFields(con,"weather")

dbListFields(con,"airports")
```

## 2. Create a data frame from a SQL pull from the database that consists only of flights that took off from JFK in May.
Select the flights table where the month is May (5) and the origin is JFK. 

```{r}
#SQL request, flight info combined with weather info
flight_text<-"SELECT f.month, f.origin, f.flight
              FROM flights f
              WHERE f.month=5 AND f.origin='JFK'"
flight_req<-dbSendQuery(con,flight_text)

flight_df<-dbFetch(flight_req,n=-1)

dbClearResult(flight_req)

head(flight_df)
```

## 3. Create a data frame from a SQL pull from the database that consists only of flights that took off on-time (a delay of less than 10 minutes) from Newark at temperatures of less than 40 degrees F.

Select the flights table where departure delay < 10 and the weather table where the temperature < 40.

```{r}
#SQL request, flight info combined with weather info
weather_text<-"SELECT f.dep_delay, f.origin, w.temp, f.flight
              FROM weather w
              JOIN flights f 
              ON f.year=w.year
              AND f.month=w.month
              AND f.day=w.day
              WHERE f.dep_delay<10 AND f.origin='EWR' AND w.temp<40"

weather_req<-dbSendQuery(con,weather_text)

weather_df<-dbFetch(weather_req,n=-1)

dbClearResult(weather_req)

head(weather_df)

```

## 4. Create data frame from a SQL pull from the database that consists of planes flown by United.
Requesting all carriers using UA (UNITED) from flight table.
```{r}
#New request: which carriers have longer delays? 
united_text<-"SELECT f.flight, f.carrier 
            FROM flights f 
            WHERE f.carrier='UA'"
            
united<-dbSendQuery(con,united_text)

united_df<-dbFetch(united,n=-1)

dbClearResult(united)

head(united_df)
```
```{r}
flights_latlon <- flights %>%
  inner_join(select(airports, origin = faa, origin_lat = lat, origin_lon = lon),
    by = "origin"
  ) %>%
  inner_join(select(airports, dest = faa, dest_lat = lat, dest_lon = lon),
    by = "dest"
  )
```

```{r}
flights_latlon %>%
  slice(1:100) %>%
  ggplot(aes(
    x = origin_lon, xend = dest_lon,
    y = origin_lat, yend = dest_lat
  )) +
  borders("state") +
  geom_segment(arrow = arrow(length = unit(0.1, "cm"))) +
  coord_quickmap() +
  labs(y = "Latitude", x = "Longitude")
```


```{r}
## Disconnect when done
dbDisconnect(con)
```
