---
title: "12-assignment_Hurst.Rmd"
author: "Jason Hurst"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## 1. Download the student evaluation dataset from the website.
The data was provided as an excel sheet in the files. I imported after installing libraries.

#Get libraries
```{r}
rm(list = ls())
library(tidyverse)
library(stats)
library(flexclust)
library(ggplot2)
library(LICORS)
library(knitr)
library(modelr)
library(readxl)
library(readxl)
turki <- read_excel("/home/brick/Dropbox/Vandy/Classes/8200/Resources/week12_resources_new_2_2_2/turkiyestudentevaluation_selected.xlsx")
View(turki)
```


## 2. Select a set of 10 related variables that you think might serve as the basis for clustering.

I chose 10 variables and kept ID data, attendance data, and perceived difficulty. The explanation can be found at the following websites: http://archive.ics.uci.edu/ml/datasets/turkiye+student+evaluation

This is the selected questions and identifiers. The unwanted columns were deleted prior to importing the excel file.

    instr: Instructor's identifier; values taken from {1,2,3} 
    class: Course code (descriptor); values taken from {1-13} 
    repeat: Number of times the student is taking this course; values taken from {0,1,2,3,...} 
    attendance: Code of the level of attendance; values from {0, 1, 2, 3, 4} 
    difficulty: Level of difficulty of the course as perceived by the student; values taken from {1,2,3,4,5} 
    Q9: enjoyment I greatly enjoyed the class and was eager to actively participate during the lectures. 
    Q13: knowledge The Instructor's knowledge was relevant and up to date. 
    Q14: prepared The Instructor came prepared for classes. 
    Q20: eager The Instructor explained the course and was eager to be helpful to students. 
    Q21: positive The Instructor demonstrated a positive approach to students. 
    Q22: respect The Instructor was open and respectful of the views of students about the course. 
    Q23: encourage The Instructor encouraged participation in the course. 
    Q25: available The Instructor responded to questions about the course inside and outside of the course. 
    Q27: solutions The Instructor provided solutions to exams and discussed them with students. 
    Q28: objective The Instructor treated all students in a right and objective manner. 

    Note: Q1-Q28 are all Likert-type, meaning that the values are taken from {1,2,3,4,5}

## 3. Run the clustering algorithm kmeanspp to identify 3 clusters.
The `kmeanspp` (stands for k-means ++) command will repeat the kmeans clustering algorithm with different starting points until it converges on a stable solution. It basically repeats the process we saw above, but with the intention of getting to a stable solution. 

```{r}
s1 <- kmeanspp(turki, k = 3, start = "random", iter.max = 1000, nstart = 50)
table(s1$cluster)
```

After obtaining 3 clusters, I practiced looking at them in regards to two variables- a student's enjoyment of the class (Spoiler cluster 2 had a high rate of students who did not enjoy the class) and whether or not the professor seemed prepared (Spoiler cluster 2 did not percieve the professor as being prepared).
```{r}
# Add predictions
turki$cluster <- s1$cluster

# Examine relationship of cluster assignment to chars.
mycols <- c("Group 1", "Group 2", "Group 3")

# Greatly enjoyed and eager to participate
pt <- prop.table(table(turki$cluster, turki$enjoyment), margin = 1)
```


```{r}
rownames(pt) <- mycols
pt <- pt * 100
pt <- round(pt, 1)
```
```{r}
kable(pt, row.names = TRUE, col.names = c("not enjoy", "little enjoy", "enjoy","much enjoy", "greatly enjoy"))
```

```{r pofessor prepared}
pt <- prop.table(table(turki$cluster, turki$prepared), margin = 1)

rownames(pt) <- mycols
pt <- pt * 100
pt <- round(pt, 1)
```
```{r}
kable(pt, row.names = TRUE, col.names = c("Not prepared", "little prepared", "prepared", "much prepared", "greatly prepared"))
```
## 4. Summarize the clusters using the variables you selected.
```{r}
# summarize groups

var.means <- colMeans(turki)

## Drop the cluster means
var.means <- var.means[-(length(var.means))]
```

# Summarizing group characteristics by cluster

The code below will summarize all of the contributing variables for each cluster, then create a plot that shows the mean of each variable within each cluster. 

First, we do our normal `summarize` command, but this time over every variable in the analysis dataset, using `summarize_all` and `funs(mean)`.

```{r}
# Summarize groups by cluster
sumt <- turki %>% group_by(cluster) %>% summarize_all(funs(mean), na.rm = TRUE)
sumt
```

Now we'll `gather` everything, so that we have just two variables: the proportion (as a number) and the variable (as a character).
```{r}
sumt2 <- gather(sumt, -cluster, key = variable, value = value)
sumt2
sumt_total <- turki %>% summarize_all(funs(mean))

sumt_total <- gather(sumt_total, key = variable, value = overall_mean)
sumt2 <- left_join(sumt2, sumt_total, by = "variable")
```

```{r}
sumt2 <- sumt2 %>% group_by(variable) %>% mutate(varmeans = mean(value))
```

## Survey Responses by Cluster
```{r}
variable_labels <- rev(c(
  "instructor ID",
  "class ID",
  "student repeat",
  "student attendance",
  "difficulty of class",
  "enjoyment",
  "professor knowledge",
  "professor prepared",
  "professor helpful",
  "professor positive",
  "professor respectful",
  "professor encouraging",
  "professor available",
  "professor discussed exams",
  "professor objective"
))

# Plot characteristics by cluster
gt <- ggplot(data = sumt2, aes(y = value, x = variable, fill = variable))
gt <- gt + geom_bar(stat = "identity") + coord_flip() + theme(legend.position = "none")
gt <- gt + scale_x_discrete(labels = variable_labels)
gt <- gt + geom_point(data = sumt2, aes(y = overall_mean, x = variable))
gt <- gt + facet_wrap(~cluster)
gt
```
