output: rmarkdown::github_document

# 1. Calculate the mean of the outcome

Load libraries

```{r  include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(Metrics)
```

## Load Dataset
United States counties data put together by the Census Bureau that summarizes the characteristics of the 3,088 counties in the United States. 

```{r data}
load("pd.Rdata")
```

## Choosing a predictive characteristic

The predictor in the example was per capita income and would be my first choice to predict home ownership. For this exercise I have chosen median household income.

```{r mean}
##Mean of median household income 2008-2012
pd%>%summarize(mean_hhinc=mean(median_hh_inc,na.rm=TRUE))
```

# 2. Create a new variable that consists of the mean of median household income


```{r create new variable in data set}
##Unconditional Average as a Predictor
pd<-pd%>%mutate(mean_hhinc=mean(median_hh_inc,na.rm=TRUE))
```

# 3. Calculate a summary measure of the errors for each observation

```{r error}
##Variable for error
pd<-pd%>%mutate(e2=median_hh_inc-mean_hhinc)

## RMSE
rmse_uncond_mean<-rmse(pd$median_hh_inc,pd$mean_hhinc)

rmse_uncond_mean

```

# 4. 

