output: rmarkdown::github_document

# 1. Calculate the mean of the outcome

Load libraries

```{r  include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(Metrics)
```

## Load Dataset
United States counties data put together by the Census Bureau that summarizes the characteristics of the 3,088 counties in the United States. 

```{r data}
load("pd.Rdata")
```

## Choosing a predictive characteristic

The predictor in the example was per capita income and would be my first choice to predict home ownership. For this exercise I have chosen median household income.

```{r mean}
##Mean of median household income 2008-2012
pd%>%summarize(mean_hhinc=mean(median_hh_inc,na.rm=TRUE))
```

# 2. Create a new variable that consists of the mean of median household income
```{r new variable}
## Create a rank variable for income 
pd<-pd%>%mutate(hhinc_rank=rank(median_hh_inc))
```

```{r create new variable in data set}
##Unconditional Average as a Predictor
pd<-pd%>%mutate(mean_hhinc=mean(median_hh_inc,na.rm=TRUE))
```

We can plot the average with the actual.

```{r}
##Plotting
gg<-ggplot(data=pd,aes(y=median_hh_inc,x=hhinc_rank,color="Actual"))
gg<-gg+geom_point(alpha=.5,size=.5)
gg<-gg+geom_point(aes(y=mean_hhinc,x=hhinc_rank,
                  color="Predicted: Unconditional Mean"),
                  size=.5)
gg<-gg+xlab("Rank of Median Household Income")+ylab("Median Household Income")
gg<-gg+scale_color_manual(name="Type",
                          values=c("Actual"="black",
                          "Predicted: Unconditional Mean"="blue")
                          )
gg<-gg+theme(legend.position="bottom")

gg

##Save for later

gg3<-gg

```

# 3. Calculate a summary measure of the errors for each observation

```{r error}
##Variable for error
pd<-pd%>%mutate(e2=median_hh_inc-mean_hhinc)

## RMSE
rmse_uncond_mean<-rmse(pd$median_hh_inc,pd$mean_hhinc)

rmse_uncond_mean

```

# 4. Calculate the mean of the outcome at levels of a predictor variable.

Although I am curious to compare ethnicity and median household income, the scope of commands is beyond my knowledge of R. For this excercise I am will be using persons below poverty.

```{r condtl_mean_single}
##Condtional Average across a single variable

## Create a variable for quartiles of college education
pd<-pd%>%mutate(poverty_level=ntile(persons_below_poverty,4))

table(pd$poverty_level)

pd<-pd%>%group_by(poverty_level)%>% ## Group by predictor
  ##Calculate mean at each level of predictor
  mutate(pred_income_poverty=mean(median_hh_inc))%>% 
  ## Ungroup
  ungroup()%>% 
  #Rank by prediction, with ties sorted randomly
  mutate(pred_income_poverty_rank=rank(pred_income_poverty,ties.method="random"))
```

# 5. Use these conditional means as a prediction: for every county, use the conditional mean to provide a ‘’best guess” as to that county’s level of the outcome.

## New Variable Home Ownership Rate
```{r}
## Create a variable for quartiles of home ownership
pd<-pd%>%mutate(homeown_rate_level=ntile(homeown_rate,4))
```

```{r}
pd%>%group_by(homeown_rate_level)%>% ## Group by predictor
  ##Calculate mean at each level of predictor
  summarise(pred_income_homeown_rate=mean(median_hh_inc))
```

```{r}
pd<-pd%>%group_by(poverty_level,homeown_rate_level)%>% ## Group by predictor
  ##Calculate mean at each level of predictor
  mutate(pred_income_poverty_and_homeown=mean(median_hh_inc))%>% 
  ## Ungroup
  ungroup()%>% 
  #Rank by prediction, with ties sorted randomly
  mutate(pred_income_poverty_and_homeown_rank=rank(pred_income_poverty_and_homeown,
                                                ties.method="random"))
```

# 6. Calculate a summary measure of the error in your predictions.

```{r}
rmse_cond_mean_two<-rmse(pd$median_hh_inc,pd$pred_income_poverty_and_homeown)
rmse_cond_mean_two
```
