output: rmarkdown::github_document

# 1. Create a graph that shows the distribution of monthly income.

Load libraries

```{r  include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(forcats)
library(RColorBrewer)
library(knitr)
```

## Load Dataset
This dataset is standard HR dataset from IBM.

```{r}
load("attrition.Rdata")
```

## Histogram of Monthly Income
```{r}
gg<-ggplot(at,aes(x=Monthly Income))
gg<-gg+geom_histogram(binwidth = 1,fill="lightblue"
gg
``` 
## Density plot of Monthly Income:
```{r}
gg<-ggplot(at,aes(x=MonthlyIncome))
gg<-gg+geom_density()
gg
```

# 2. Create a graph that shows the average level of monthly income by field of education.
```{r}

at_sum<-at%>%
  group_by(EducationField)%>%
  summarize(income_avg=mean(MonthlyIncome))

at_sum

```
## Bar Plot with aesthetics: mean Monthly Income as height, Education Field as cateogry

```{r}
gg<-ggplot(at_sum,aes(x=EducationField,y=income_avg))
## Use bar plot geometry, height of bars set by level observed in dataset
gg<-gg+geom_bar(stat="Identity")
## Print
gg
```

## Reorder by level
```{r}
gg<-ggplot(at_sum,aes(x=fct_reorder(EducationField,income_avg),y=income_avg))

gg<-gg+geom_bar(stat="identity")

## Labeling
gg<-gg+xlab("Education Field")+ylab("Monthly Income")
##Print
gg
```

# 3. Create another graph that shows average level of monthly income by field of education and gender.

## Summarize Monthly Income by field of education and gender
```{r}
at_sum<-at%>%
  group_by(EducationField,Gender)%>%
  summarize(income_avg=mean(MonthlyIncome))

## Get the results
at_sum
```

## Bar Plot of average level of monthly income by field of education and gender
```{r}
gg<-ggplot(at_sum,aes(x=fct_reorder(EducationField,income_avg),y=income_avg,color=Gender))
gg<-gg+geom_bar(stat="identity",aes(fill=Gender),position="dodge")
gg<-gg+ylab("Average Monthly Income")+xlab("Field of Education")
gg
```

# 4. Create a graph that shows average levels of monthly income by field of education, gender and job level (scale of 1-5, highest ranked employees are 5)

## Summary
```{r}

at_sum<-at%>%
  group_by(EducationField,Gender,JobLevel)%>%
  summarize(income_avg=mean(MonthlyIncome))%>%
  arrange(-income_avg)

at_sum
```

## Bar plot, with unstacked (dodge)
```{r}
gg<-ggplot(at_sum,aes(x=reorder(EducationField,income_avg),
                      y=income_avg,
                      fill=Gender))
 gg<-gg+geom_bar(stat="identity",position="dodge")
gg<-gg+facet_wrap(~JobLevel)
gg<-gg+coord_flip()
gg
```
