Use gather function week 4 #1
---
title: Wrangling Data Assignment 4
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
```{r} 
library(tidyverse)
library(haven)
library(readxl)
```

# 1. Panel data for OECD countries. Save as file name oecd.RData http://www.wiley.com/legacy/wileychi/baltagi/supp/Gasoline.dat

## Panel data for OECD countries. Save as file name oecd.RData 
```{r}
gas<-read_delim("http://www.wiley.com/legacy/wileychi/baltagi/supp/Gasoline.dat",
                delim=" ",trim_ws = TRUE)

save(gas,file="OECD.Rdata")

```
## Note that data follows Tidy principles: 
## 1. Each variable forms a column.
## 2. Each observation forms a row
## 3. Each type of observational unit forms a table.

```{r}
is.data.frame(gas)
is.tibble(gas)
is_tibble(gas)
typeof(gas)
```

## So you can do fun things with it like saving it in a variable:

```{r}
desc_gas <- gas %>%
  group_by(COUNTRY) %>%
  summarise(mean_gaspcar = mean(LGASPCAR),
            sd_gaspcar = sd(LGASPCAR),
            max_gaspcar = max(LGASPCAR),
            min_gaspcar = min(LGASPCAR))
```
## Further analysis like in this example: https://forecasters.org/wp-content/uploads/gravity_forms/7-c6dd08fee7f0065037affb5b74fec20a/2017/07/SELUKAR_RAJESH_ISF2017.pdf

## Or answers questions like: Which country has the maximum average gasoline consumption?
## Source https://b-rodrigues.github.io/modern_R/descriptive-statistics-and-data-manipulation.html

```{r}
desc_gas %>%
  filter(max(mean_gaspcar) == mean_gaspcar)
```

## Which country has the minimum consumption?

```{r}
desc_gas %>%
  filter(min(mean_gaspcar) == mean_gaspcar) 
```

# 2. U.S. National Election Survey data from 2004. Save as file name nes.RData. http://www.uta.edu/faculty/story/DataSets.htm

## U.S. National Election Survey data from 2004. Save as file name nes.RData. 
```{r}
nes<-read_spss("http://www.uta.edu/faculty/story/DataSets/NES2004.sav")
# Now need to clean up 
# Get rid of unwanted columns
nes <-nes [ ,-(c(3))]
# save file
save(nes,file="nes.Rdata")
```

# 3. General social survey, student version. Save as file name gss.Rdata. http://www.uta.edu/faculty/story/DataSets.htm

## General social survey, student version. Save as file name gss.Rdata. 
```{r}
if(file.exists("gss.xls")==FALSE)
{download.file("http://www.uta.edu/faculty/story/DataSets/GeneralSocialSurvey1996ExcelDataStudentVersion.xls", destfile = "gss.xls")
gss<-read_excel("gss.xls",skip=0,col_names=TRUE)
}else{gss<-read_excel("gss.xls",skip=0,col_names=TRUE)}
```
## Check data
```{r verify tibble}
is.data.frame(gss)
is.tibble(gss)
is_tibble(gss)
typeof(gss)
```
## Save file
```{r save file}
save(gss,file="gss.Rdata")
```

# 4. Replication file for “STATISTICAL DISCRIMINATION OR PREJUDICE? A LARGE SAMPLE FIELD EXPERIMENT”. Open up and save the mainData.csv file. Save it as maindata.RData. Available here, look under the “Data and Analysis” tab.
## Web page: http://thedata.harvard.edu/dvn/dv/restat/faces/study/StudyPage.xhtml?globalId=doi:10.7910/DVN/26410&studyListingIndex=1_f1a615c2cbee97fb79c9d575ae73

```{r}
#Load in the mainData dataset
library(readr)
mainData<-read_csv(file="https://dataverse.harvard.edu/api/access/datafile/:persistentId?persistentId=doi:10.7910/DVN/26410/05NHUX")
write_csv(mainData,path="mainData.csv")

# Get rid of unwanted columns
mainData <-mainData [ ,-(c(1))]

#Check it out
head(mainData)
```

```{r verify tibble}
is.data.frame(mainData)
is.tibble(mainData)
is_tibble(mainData)
typeof(mainData)
```
```{r}
#Saving as an R file
save(mainData,file="mainData.Rdata")
```

# 5. The Lalonde dataset, covering work experiences in the Panel Study of Income Dynamics (psid). Save as psid.RData http://users.nber.org/~rdehejia/data/psid_controls.txt

## The variables from left to right are: treatment indicator (1 if treated, 0 if not treated), age, education, Black (1 if black, 0 otherwise), Hispanic (1 if Hispanic, 0 otherwise), married (1 if married, 0 otherwise), nodegree (1 if no degree, 0 otherwise), RE74 (earnings in 1974), RE75 (earnings in 1975), and RE78 (earnings in 1978). 
```{r}
psid_controls <- read.table("http://www.nber.org/~rdehejia/data/psid_controls.txt", quote="\"", comment.char="")
```
## Label Data
```{r PSID labels}
names(psid_controls)<-c("treatment",
                "age",
                "education",
                "black",
                "hispanic",
                "married",
                "nodegree",
                "earnings1974",
                "earnings1975",
                "earnings1978")
head(psid_controls)
```


```{r Save dataset as Rdata file}
save(psid_controls,file="psid.Rdata")
```
