---
title: "Assignment-7"
author: "Jason Hurst"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ways of getting data from the web

We will cover several different ways you can get data from the web

1. Directly downloading web pages via the `url()` command. 
1. Scraping simple web tables via `read_html()` and `html_table()` command
1. Interacting with Application Programming Interfaces (APIs) via R libraries that have been designed as "wrappers" for these interfaces, like the awesome `acs` library and the `tigris` library for geographic shapes. 
1. Interacting with APIs directly, 


## Libraries

We will use multiple new libraries today. Among the ones you'll need: 

* `rvest` for scraping websites

* `acs` for accessing American Community Survey data via the census API


```{r}
library(tidyverse)
library(rvest)
library(acs)
library(lubridate)
library(noncensus)
library(tigris)
```
## 1. Download data for all of the census tracts in Fulton County, GA on education levels.

# First obtain Zip Code Level Data from the American Community Survey

#We start by using the lookup_code from the `tigris` package to get the fips codes Fulton County, GA.
```{r}
## Look up fips code for county
lookup_code("GA","Fulton") 

state_fips<-"13"
county_stub<-"121"
```

Next, we'll combine the state and county fips into a single object

```{r}
county_fips<-paste0(state_fips,county_stub)
```

```{r}
# Get dataset that matches all zip codes to cities, counties and states. 
county_to_zip<-read_csv("http://www2.census.gov/geo/docs/maps-data/data/rel/zcta_county_rel_10.txt")
save(county_to_zip,file="county_to_zip.Rdata")

#easier names to work with
names(county_to_zip)<-tolower(names(county_to_zip))

#Just zip codes in selected county
county_to_zip<-county_to_zip%>%
  filter(state==as.numeric(state_fips),county==county_stub)%>%
  select(zcta5,state,county)

#list of zip codes
ziplist<-county_to_zip$zcta5

#City names
data(zip_codes)

city_zip<-zip_codes%>%filter(zip%in%ziplist)%>%select(zip,city)

#Arrange in order
city_zip<-city_zip%>%arrange(as.numeric(zip))
```
# Use key to access data

```{r}
# load in ACS Key
my_acs_key<-readLines("/home/brick/Dropbox/Vandy/Classes/8210/CensusAPI.txt",warn = FALSE)
acs_key<-my_acs_key

# Or just paste it here.
#acs_key<-"<Nice Try!>"

#List of tables: https://www.census.gov/programs-surveys/acs/technical-documentation/summary-file-documentation.html under, 1-year appendices
# We will use this one- b15002: education of pop over 25, by sex 
# Another is b19001: household income over last 12 months

api.key.install(acs_key, file = "key.rda")

select_zip<-geo.make(zip.code=ziplist)

county_educ=acs.fetch(geography=select_zip,
                      endyear=2016,
                      table.number="B15002",
                      col.names="pretty",verbose=T)
#save(county_educ,file="county_educ_la.Rdata")  # <---- this may take a long time :(
acs.colnames(county_educ)
```

## 2. Compute the proportion of the population that has a bachelor’s degree or above by census tract.

```{r}
## Proportion of individuals at college or above=
## number with college degree/
## total number
bachpct<-divide.acs(numerator=(county_educ[,15]+
                                      county_educ[,16]+
                                      county_educ[,17]+
                                      county_educ[,18]+
                                      county_educ[,32]+
                                      county_educ[,33]+
                                      county_educ[,34]+
                                      county_educ[,35]),
                            denominator=county_educ[,1])
### Double check -- 1 is total population and others are groups with Bachelor's or above. check with acs.colnames(county_educ)

head(bachpct)
```

## 3. Download data for all of the zip codes in LA county on family income by census tract.

```{r}
## Look up fips code for county
lookup_code("CA","Los Angeles") 

lastate_fips<-"06"
lacounty_stub<-"037"
```
Next, we'll combine the state and county fips into a single object

```{r}
lacounty_fips<-paste0(lastate_fips,lacounty_stub)
```

```{r}
#Redownload the dataset that matches all zip codes to cities, counties and states. Note that some steps could be skipped if we use different nomenclature for getting data about GA
lacounty_to_zip<-read_csv("http://www2.census.gov/geo/docs/maps-data/data/rel/zcta_county_rel_10.txt")
save(lacounty_to_zip,file="lacounty_to_zip.Rdata")

#easier names to work with
names(lacounty_to_zip)<-tolower(names(lacounty_to_zip))

#Just zip codes in selected county
lacounty_to_zip<-lacounty_to_zip%>%
  filter(state==as.numeric(lastate_fips),county==lacounty_stub)%>%
  select(zcta5,state,county)

#list of zip codes
laziplist<-lacounty_to_zip$zcta5

#City names
data(zip_codes)

lacity_zip<-zip_codes%>%filter(zip%in%laziplist)%>%select(zip,city)

#Arrange in order
lacity_zip<-lacity_zip%>%arrange(as.numeric(zip))
```

#List of tables: https://www.census.gov/programs-surveys/acs/technical-documentation/summary-file-documentation.html under, 1-year appendices
# Now we will use this one- b19001: household income over last 12 months

## Family Income Data
```{r}

# 19001-- family income     

laselect_zip<-geo.make(zip.code=laziplist)
lacounty_income<-acs.fetch(geography=laselect_zip, 
                        endyear = 2016,
                        table.number="B19001", 
                        col.names="pretty")

acs.colnames(lacounty_income)

head(lacounty_income)
```
 
## 4. Compute the proportion of the population that has family income above 75,000 in each census tract.

```{r}
#Proportion above 75k-- 
prop_above_75<-divide.acs(numerator=(lacounty_income[,13]+
                            lacounty_income[,14]+
                            lacounty_income[,15]+
                            lacounty_income[,16]+
                            lacounty_income[,17]),
                          denominator=lacounty_income[,1]
                          )
head(prop_above_75)
```

## 5. Plot the proportion of residents in each census tract with incomes above 75,000 as a function of income.

```{r}
county_income2<-acs.fetch(geography=laselect_zip, 
                        endyear = 2016,
                        table.number="B19001", 
                        col.names="pretty")


head(county_income2)
```

```{r}
# Convert to tibble
lacounty_df<-tibble(substr(geography(lacounty_income)[[1]],7,11),
                    as.numeric(estimate(county_income2)),
                    as.numeric(estimate(prop_above_75))
)


# Give it easy to use names
names(lacounty_df)<-c("zip", "income", "income_75")
save(lacounty_df,file="la.RData")

head(lacounty_df)
```

```{r}

gg<-ggplot(lacounty_df,aes(x=county_income2,y=income_75))
gg<-gg+geom_point()
gg

```
