---
title: "Assignment-7"
author: "Jason Hurst"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ways of getting data from the web

1. Directly downloading web pages via the `url()` command. 
2. Scraping simple web tables via `read_html()` and `html_table()` command
3. Interacting with Application Programming Interfaces (APIs) via R libraries that have been designed as "wrappers" for these interfaces, like the awesome `acs` library and the `tigris` library for geographic shapes. 
4. Interacting with APIs directly, 


## Libraries

We will use multiple new libraries today. Among the ones you'll need: 

* `rvest` for scraping websites

* `acs` for accessing American Community Survey data via the census API


```{r}
library(tidyverse)
library(rvest)
library(acs)
library(lubridate)
library(noncensus)
library(tigris)
library(tidycensus)
```

## 1. Download data for all of the zip codes in Los Angeles county on education levels.
## Zip Code Level Data from the American Community Survey

The first step is to create a list of all zip codes in Los Angeles County. We can do this by using another dataset that includes a comprehensive listing of zip codes by county and city. 

We start by using the lookup_code from the `tigris` package to get the fips codes for Los Angeles County in CA.

```{r}
## Look up fips code for county
lookup_code("CA","Los Angeles") 

lastate_fips<-"06"
lacounty_stub<-"037"
```

Next, we'll combine the state and county fips into a single object

```{r}
lacounty_fips<-paste0(lastate_fips,lacounty_stub)
```

```{r}
# Get dataset that matches all zip codes to cities, counties and states. 
lacounty_to_zip<-read_csv("http://www2.census.gov/geo/docs/maps-data/data/rel/zcta_county_rel_10.txt")
save(lacounty_to_zip,file="lacounty_to_zip.Rdata")

#easier names to work with
names(lacounty_to_zip)<-tolower(names(lacounty_to_zip))

#Just zip codes in selected county
lacounty_to_zip<-lacounty_to_zip%>%
  filter(state==as.numeric(lastate_fips),county==lacounty_stub)%>%
  select(zcta5,state,county)

#list of zip codes
laziplist<-lacounty_to_zip$zcta5

#City names
data(zip_codes)

lacity_zip<-zip_codes%>%filter(zip%in%laziplist)%>%select(zip,city)

#Arrange in order
lacity_zip<-lacity_zip%>%arrange(as.numeric(zip))
```

# Use key to access data

```{r}
# load in ACS Key
my_acs_key<-readLines("/home/brick/Dropbox/Vandy/Classes/8210/CensusAPI.txt",warn = FALSE)
acs_key<-my_acs_key

# Or just paste it here.
#acs_key<-"<Nice Try!>"

#List of tables: https://www.census.gov/programs-surveys/acs/technical-documentation/summary-file-documentation.html under, 1-year appendices
# We will use this one- b15002: education of pop over 25, by sex 
# Another is b19001: household income over last 12 months

api.key.install(acs_key, file = "key.rda")

laselect_zip<-geo.make(zip.code=laziplist)

lacounty_educ=acs.fetch(geography=laselect_zip,
                      endyear=2016,
                      table.number="B15002",
                      col.names="pretty",verbose=T)
#save(lacounty_educ,file="lacounty_educ_la.Rdata")  # <---- this may take a long time :(
acs.colnames(lacounty_educ)
```
## 2. Compute the proportion of the population that has a bachelor’s degree or above by census tract.

```{r}
## Proportion of individuals at college or above=
## number with college degree/
## total number
laprop_coll_above<-divide.acs(numerator=(lacounty_educ[,15]+
                                      lacounty_educ[,16]+
                                      lacounty_educ[,17]+
                                      lacounty_educ[,18]+
                                      lacounty_educ[,32]+
                                      lacounty_educ[,33]+
                                      lacounty_educ[,34]+
                                      lacounty_educ[,35]),
                            denominator=lacounty_educ[,1])
### Double check -- 1 is total population and others are groups with Bachelor's or above. check with acs.colnames(lacounty_educ)

head(laprop_coll_above)
```

## 3. Download data for all of the zip codes in LA county on family income by census tract.
#List of tables: https://www.census.gov/programs-surveys/acs/technical-documentation/summary-file-documentation.html under, 1-year appendices
# Now we will use this one- b19001: household income over last 12 months
## Family Income Data
```{r}

# 19001-- family income     

laselect_zip<-geo.make(zip.code=laziplist)
lacounty_income<-acs.fetch(geography=laselect_zip, 
                        endyear = 2016,
                        table.number="B19001", 
                        col.names="pretty")

acs.colnames(lacounty_income)

head(lacounty_income)
```

## 4. Compute the proportion of the population that has family income above 75,000 in each census tract.

```{r}
#Proportion above 75k-- 
prop_above_75<-divide.acs(numerator=(lacounty_income[,13]+
                            lacounty_income[,14]+
                            lacounty_income[,15]+
                            lacounty_income[,16]+
                            lacounty_income[,17]),
                          denominator=lacounty_income[,1]
                          )
head(prop_above_75)
```


## 5. Plot the proportion of residents with incomes above 75,000 as a function of income.
```{r}
# Convert to tibble
lacounty_df<-tibble(substr(geography(lacounty_educ)[[1]],7,11),
                       as.numeric(estimate(laprop_coll_above)),
                       as.numeric(estimate(prop_above_75))
)


# Give it easy to use names
names(lacounty_df)<-c("zip","college_educ","income_75")
save(lacounty_df,file="dav.RData")

head(lacounty_df)
```

## Get income data for zip codes
```{r}
#income total-- 
census_api_key ("Your Census Key")
lacounty_income2<- get_acs(geography = "zcta",
                           year= 2016 ,
                           variables = "B19001_001")
## Clean up data

wordstoremove <- c("ZCTA5")
(dat1 <- as.data.frame(sapply(lacounty_income2, function(x) 
    gsub(paste(wordstoremove, collapse = '|'), '', x))))
 dat2 <- dat1[,c(2,4,5)]
names(dat2) <- c("zip","estimate", "moe")

```
## Combine by zip code with LA County zip codes to filter LA County from country
```{r}
dat2$zip <- as.numeric(dat2$zip)
lacity_zip$zip <- as.numeric(lacity_zip$zip)
dat3<-left_join(lacity_zip,dat2,by=c("zip"))
```
## Combine dat3 with lacounty_df
```{r}

lacounty_df$zip <- as.numeric(lacounty_df$zip)
dat4<-left_join(lacounty_df,dat3,by=c("zip"))
```

## Graph proportion of residents with incomes above 75,000 as a function of income
```{r}

gg<-ggplot(dat4,aes(x=estimate,y=income_75))
gg<-gg+geom_point()
gg

```


## Bonus: Compare LA County to the rest of California
```{r}
census_api_key ("your census key")
ca <- get_acs(geography = "county", 
              year= 2016 ,
              variables = c(income = "B19001_001"), 
              state = "CA")
```

```{r}
ca %>%
  mutate(NAME = gsub(" County, California", "", NAME)) %>%
  ggplot(aes(x = estimate, y = reorder(NAME, estimate))) +
  geom_errorbarh(aes(xmin = estimate - moe, xmax = estimate + moe)) +
  geom_point(color = "red", size = 3) +
  labs(title = "Household income by county in California",
       subtitle = "2016 American Community Survey",
       y = "",
       x = "ACS estimate (bars represent margin of error)")
```
