---
title: "Assignment-9"
author: "Jason Hurst"
output: github_document
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(modelr)
library(knitr)
library(groupdata2)
load("training.RData")
# For this assignment, use the lemons dataset, which is a subset of the dataset used for a Kaggle competition described here: https://www.kaggle.com/c/DontGetKicked/data.
```

## 1. Using the lemons dataset, plot the probability of a car being a bad buy by make.

We always want to start with a cross tab of our dependent variable as a function of other variables. We structure cross tabs in a very particular way for the purposes of research: the independent variable goes on the rows, and the dependent variable goes on the columns. If proportions or percentages are going to be calculated, they should be calculated across rows. 

Let's look at a crosstab of `IsBadBuy` with the independent variable of `Make`

```{r}
tab_make<-with(training,table(Make,IsBadBuy))
```

Change the row and column titles

```{r}
colnames(tab_make)<-c("Not a Lemon","Lemon")
kable(tab_make)
```
Add proportions to this table.
```{r}
tab_make_prop<-prop.table(tab_make,margin=1)
kable(tab_make_prop)
```

Audiences prefer percentages. Easy enough to do:
```{r}
kable(round(tab_make_prop*100,1))
```

To format the data for barcharts, we make use of the concept of conditional means. Let's use the variable "Make" to calculate the probability of the car being a lemon. 

```{r}
training_sum <- training%>%
  group_by(Make)%>%
  summarise(prob_bb=mean(IsBadBuy,na.rm=T))
```

Plot with ggplot 

```{r}
gg1 <- ggplot(training_sum, aes(y=prob_bb, x=Make))+
  geom_bar(stat = "identity", position = "dodge")+
  labs(title = "Probability of Lemons", x="Make of Vehicle", y="Lemon")+
  theme(legend.title = element_blank())
gg1 <- gg1+geom_text(aes(label=round(prob_bb,2)),
                     position = position_dodge(width = .9),
                     vjust=.25)
gg1
```

## 2. Create a table that shows the probability of a car being a bad buy by make.

```{r}
prop.table(tab_make,margin=1)%>%kable()
```

## 3. Create a heatmap of the probability of a car being a bad buy by make and acquisition type.
## Note AcquisitionType is a category listed in the Carvana_Data_Dictionary.txt but doesn't exist in the data (training or test). I will use Color instead.
To generate a heat map, we'll first divide up the independent variables into quintiles:

```{r}
training <- training%>%
  mutate(Make_quintile=ntile(Make,5),
         Color_quintile=ntile(Color,5))
```

Then we'll create a summary dataset that shows the probabilitie of the outcome across all of the combined categories of the two independent variables. 

```{R}
training_sum<-training%>%
  group_by(Make_quintile,Color_quintile)%>%
  summarize(prob_make=mean(IsBadBuy,na.rm=TRUE))%>%
  arrange(-prob_make)
```

Missing data isn't important, so we'll drop it. 

```{r}
trainin_sum<-training_sum%>%filter(!(is.na(Make_quintile)),!(is.na(Color_quintile)))
```

Now we're ready to plot!

```{r}
gg<-ggplot(training_sum,
           aes(x=as.factor(Make_quintile),
               y=as.factor(Color_quintile),fill=prob_make))
gg<-gg+geom_tile()
gg<-gg+scale_fill_gradient(low="white",high="red")
gg<-gg+xlab("Make")+ylab("Color")
gg<-gg+theme(legend.title=element_blank())
gg
```


## 4. Create a plot of your choosing that shows the probability of a car being a bad buy by year and make.

```{r}
tab_make_year<-with(filter(training,
                           as.character(VehYear)!="Year of Car"),
                           table(Make,IsBadBuy))
```

*Outcome by "Make" AND "Year"*
```{r}
prop.table(tab_make_year,margin=1)%>%kable()
```

## Bar Graphs from Cross Tabs

To format the data for barcharts, we make use of the concept of conditional means. Let's use two variables to calculate the probability of receiving a pizza: poor and student. 

```{r}
training2_sum<-training%>%
  group_by(VehYear,Make)%>%
  summarize(prob_lemon=mean(IsBadBuy,na.rm=TRUE))
```

Then we can plot this using our familiar ggplot commands:

```{r}
gg2<-ggplot(training2_sum,aes(y=prob_lemon,x=Make,fill=VehYear))
gg2<-gg2+geom_bar(stat="identity",position="dodge")
gg2<-gg2+xlab("Make")+ylab("Pr(Lemon)")
gg2<-gg2+theme(legend.title=element_blank())

gg2<-gg2+geom_text(aes(label=round(prob_lemon,2)),
                   position=position_dodge(width=.9),
                   vjust=-.25)
gg2
```


